
## Studies stopped because of safety

### Safety by Therapeutic area

```{r, safety_byTA, cache = TRUE}
p_safety_by_ta <- all_tests %>%
    filter(predictionType == "metareason") %>%
    filter(prediction == "Safety_Sideeffects") %>%
    # filter(predictionType %in% c("reason", "clinical", "stopStatus")) %>%
    filter(comparisonType == "ta") %>%
    mutate(comparison = fct_reorder(comparison, estimate, median)) %>%
    ggplot(aes(x = comparison, y = estimate)) +
    geom_hline(aes(yintercept = 1),
        size = .25,
        linetype = "dashed"
    ) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
        width = 0
    ) +
    geom_point() +
    scale_y_log10(name = "Odds Ratio") +
    coord_flip() +
    facet_grid(. ~ prediction) +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        axis.title.y = element_blank()
    )
p_safety_by_ta
```

```{r, safety_main, cache = TRUE}
prepare_safety_data <- function(tests) {
    tests %>%
        filter(predictionType == "metareason") %>%
        filter(prediction == "Safety_Sideeffects") %>%
        filter(comparisonType %in% c(
            "ta",
            "geneticConstraint",
            "rnaSpecificity",
            "lof_tolerance",
            "interactions"
        )) %>%
        filter(comparison != "Not detected") %>%
        mutate(comparison = fct_reorder(comparison, estimate)) %>%
        mutate(comparison = fct_relevel(
            comparison,
            "Tissue enriched",
            "Group enriched",
            "Tissue enhanced",
            "Low tissue specificity"
        )) %>%
        mutate(comparison = fct_relevel(
            comparison,
            "greaterThan20",
            after = Inf
        )) %>%
        mutate(comparison = fct_recode(comparison,
            "Very high" = "0",
            "High" = "1",
            "Medium" = "2",
            "Low" = "3",
            "Very low" = "4",
            "Lowest" = "5"
        )) %>%
        mutate(comparison = fct_recode(comparison,
            ">20 partners" = "greaterThan20",
            "11-20 partners" = "from11to20",
            "1-10 partners" = "from1to10",
            "None reported" = "none"
        )) %>%
        mutate(comparison = fct_recode(comparison,
            "Intolerant (pLI > 0.9)" = "LoF intolerant",
            "Tolerant (pLI < 0.1)" = "LoF tolerant"
        )) %>%
        mutate(comparisonType = fct_recode(
            comparisonType,
            "LoF" = "lof_tolerance",
            "Interacting proteins" = "interactions",
            "RNA specificity" = "rnaSpecificity",
            "Indication" = "ta",
            "Genetic Constraint (LOEUF bin)" = "geneticConstraint"
        )) %>%
        mutate(comparisonType = fct_rev(
            fct_reorder(comparisonType, estimate, max, na.rm = TRUE)
        )) %>%
        mutate(comparisonType = fct_relevel(
            comparisonType, "Interacting proteins",
            after = Inf
        )) %>%
        mutate(oddsinterval = paste(round(estimate, digits = 3),
            " (",
            round(conf.low, digits = 2),
            "-",
            round(conf.high, digits = 2),
            ")",
            sep = ""
        )) %>%
        mutate(significance = format_pval(p.value))
}

plot_safety_dots <- function(toplot_data) {
    toplot_data %>%
        ggplot(aes(x = comparison, y = estimate)) +
        geom_hline(aes(yintercept = 1),
            size = .25,
            linetype = "dashed"
        ) +
        geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
            width = 0, color = "steelblue", size = 1
        ) +
        geom_point(color = "steelblue", size = 3) +
        scale_y_log10(name = "Odds Ratio", position = "right") +
        scale_x_discrete(labels = function(x) {
            stringr::str_wrap(x, width = 25)
        }) +
        coord_flip() +
        facet_grid(comparisonType ~ .,
            scale = "free_y",
            space = "free_y",
            switch = "y"
        ) +
        theme_cowplot(font_size = 12) +
        panel_border(color = "black") +
        theme(
            axis.title.y = element_blank(),
            axis.line = element_blank(),
            panel.spacing = unit(-0.1, "lines"),
            strip.placement = "outside",
            strip.background = element_blank(),
            strip.switch.pad.grid = unit(20, "pt")
        )
}

plot_safety_numbers <- function(toplot_data) {
    toplot_data %>%
        mutate(a = formatC(a, format = "d", big.mark = ",")) %>%
        select(
            comparison,
            comparisonType,
            prediction,
            a,
            oddsinterval,
            significance
        ) %>%
        gather(key, value, -comparison, -comparisonType, -prediction) %>%
        mutate(key = fct_recode(key,
            "n" = "a",
            "OR (95% CI)" = "oddsinterval",
            "P-value" = "significance"
        )) %>%
        ggplot(aes(x = comparison, y = key)) +
        # geom_tile(fill = NA, color = "black") +
        geom_tile(fill = NA) +
        geom_text(aes(label = value)) +
        scale_x_discrete(labels = function(x) stringr::str_wrap(x, width = 25)) +
        scale_y_discrete(expand = c(0, 0), position = "right") +
        coord_flip() +
        facet_grid(comparisonType ~ .,
            scale = "free_y",
            space = "free_y"
        ) +
        theme_cowplot(font_size = 12) +
        panel_border(color = "black") +
        theme(
            axis.ticks = element_blank(),
            axis.title = element_blank(),
            axis.text.y = element_blank(),
            panel.spacing = unit(-0.1, "lines"),
            strip.background = element_blank(),
            axis.line = element_blank(),
            strip.text.y = element_blank()
        )
}

toplot_safety_data <- prepare_safety_data(all_tests)
p_safety_main_dots <- plot_safety_dots(toplot_safety_data)
p_safety_main_numbers <- plot_safety_numbers(toplot_safety_data)

p_safety_main <- plot_grid(
    p_safety_main_dots +
        theme(plot.margin = margin(
            t = 5.5, b = 5.5, l = 5.5, r = -4, unit = "pt"
        )),
    p_safety_main_numbers +
        theme(plot.margin = margin(
            t = 5.5, b = 5.5, l = 0, r = 5.5, unit = "pt"
        )),
    rel_widths = c(1, 0.65),
    align = "h",
    axis = "t",
    nrow = 1,
    ncol = 2
)
p_safety_main
```

### Safety by somatic T-D evidence

```{r, safety_by_cancer_datasource, cache = TRUE}
cancer_ds <- c("cancer_gene_census", "intogen", "eva_somatic")
p_safety_by_cancer_datasource <- all_tests %>%
    filter(prediction != "Success") %>%
    filter(predictionType %in% c("metareason", "clinical", "stopStatus")) %>%
    filter(comparison %in% cancer_ds) %>%
    filter(comparisonType == "byDatasource") %>%
    mutate(predictionType = fct_relevel(
        predictionType,
        "clinical", "stopStatus", "metareason"
    )) %>%
    mutate(predictionType = fct_recode(
        predictionType,
        "All studies" = "clinical",
        "Stopped studies" = "stopStatus",
        "Reason implication" = "metareason"
    )) %>%
    mutate(prediction = fct_recode(prediction,
        "Suspicious" = "Business_Administrative",
        "Invalid" = "Invalid_Reason",
        "Safety" = "Safety_Sideeffects",
        "Negative" = "Negative",
        "Neutral" = "Neutral"
    )) %>%
    mutate(prediction = fct_reorder(prediction, estimate, median)) %>%
    mutate(prediction = fct_relevel(prediction, "Phase III+", after = Inf)) %>%
    mutate(prediction = fct_relevel(prediction, "Phase IV", after = Inf)) %>%
    mutate(comparison = fct_recode(
        comparison,
        "ClinVar (Somatic)" = "eva_somatic",
        "IntOgen" = "intogen",
        "COSMIC (CGC)" = "cancer_gene_census"
    )) %>%
    ggplot(aes(x = prediction, y = estimate)) +
    geom_hline(aes(yintercept = 1),
        size = .25,
        linetype = "dashed"
    ) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
        width = 0
    ) +
    geom_point() +
    scale_y_log10(name = "Odds Ratio") +
    coord_flip() +
    facet_grid(
        predictionType ~ comparison,
        scales = "free_y",
        space = "free_y"
    ) +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        axis.title.y = element_blank(),
        strip.background = element_rect(fill = NA),
        panel.grid.major.y = element_blank()
    )
p_safety_by_cancer_datasource
```

### Safety by target genetic constraint

```{r, safety_by_genetic_constrain, cache = TRUE}
p_safety_by_genetic_constrain <- all_tests %>%
    filter(comparisonType == "geneticConstraint") %>%
    filter(predictionType == "metareason") %>%
    filter(prediction == "Safety_Sideeffects") %>%
    ggplot(aes(
        x = comparison,
        y = estimate,
        color = prediction
    )) +
    geom_hline(aes(yintercept = 1),
        size = .25,
        linetype = "dashed"
    ) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
        width = 0
    ) +
    geom_point() +
    scale_color_discrete(name = "Study stop reason") +
    scale_y_log10(name = "Odds Ratio") +
    scale_x_discrete(name = "pLOEUF target genetic constraint bin (GnomAD)") +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        legend.position = c(0.95, 0.95),
        legend.justification = c(1, 1),
    )
p_safety_by_genetic_constrain
```

### RNA specificity

- Tissue Enriched: Genes with an expression level greater than 1 (TPM or FPKM) that also have at least five-fold higher expression levels in a particular tissue compared to all other tissues.
- Group Enriched: Genes with an expression level greater than 1 (TPM or FPKM) that also have at least five-fold higher expression levels in a group of 2-7 tissues compared to all other tissues, and that are not considered Tissue Enriched.
- Tissue Enhanced: Genes with an expression level greater than 1 (TPM or FPKM) that also have at least five-fold higher expression levels in a particular tissue compared to the average levels in all other tissues, and that are not considered Tissue Enriched or Group Enriched.

```{r, safety_rna_specificity, cache = TRUE}
p_safety_rna_specificity <- all_tests %>%
    filter(predictionType == "metareason") %>%
    filter(prediction == "Safety_Sideeffects") %>%
    filter(comparisonType == "rnaSpecificity") %>%
    filter(comparison != "Not detected") %>%
    mutate(comparison = fct_relevel(
        comparison,
        "Low tissue specificity",
        "Tissue enhanced",
        "Group enriched",
        "Tissue enriched"
    )) %>%
    ggplot(aes(
        x = comparison,
        y = estimate,
        color = prediction
    )) +
    geom_hline(aes(yintercept = 1),
        size = .25,
        linetype = "dashed"
    ) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high),
        width = 0
    ) +
    geom_point() +
    scale_color_discrete(name = "Study stop reason") +
    scale_y_log10(name = "Odds Ratio") +
    scale_x_discrete(name = "RNA tissue specificity") +
    theme_cowplot(font_size = 12) +
    panel_border(color = "black") +
    theme(
        legend.position = c(0.95, 0.95),
        legend.justification = c(1, 1),
    )
p_safety_rna_specificity
```

## All vs oncology vs non-oncology

```{r, safety_all_vs_oncology, cache = TRUE}
# p_safety_by_indication <- prepare_safety_data(indication_filtered_tests) %>%
#     filter(comparisonType != "Indication") %>%
#     ggplot(aes(x = comparison, y = estimate)) +
#     geom_hline(aes(yintercept = 1),
#         linewidth = .25,
#         linetype = "dashed"
#     ) +
#     geom_errorbar(
#         aes(
#             ymin = conf.low,
#             ymax = conf.high,
#             color = indication_class
#         ),
#         position = position_dodge(width = 0.8),
#         width = 0,
#         size = 1
#     ) +
#     geom_point(aes(color = indication_class),
#         size = 3,
#         position = position_dodge(width = 0.8)
#     ) +
#     scale_y_log10(name = "Odds Ratio", position = "right") +
#     scale_x_discrete(labels = function(x) {
#         stringr::str_wrap(x, width = 25)
#     }) +
#     coord_flip() +
#     facet_grid(comparisonType ~ .,
#         scale = "free_y",
#         space = "free_y",
#         switch = "y"
#     ) +
#     theme_cowplot(font_size = 12) +
#     scale_color_brewer(palette = "Set1", name = "Indication") +
#     panel_border(color = "black") +
#     theme(
#         axis.title.y = element_blank(),
#         axis.line = element_blank(),
#         panel.spacing = unit(-0.1, "lines"),
#         strip.placement = "outside",
#         strip.background = element_blank(),
#         strip.switch.pad.grid = unit(20, "pt"),
#         legend.position = c(0.05, 0.95),
#     )
# p_safety_by_indication

toplot_safety_data_byindication <- prepare_safety_data(indication_filtered_tests) %>% filter(comparisonType != "Indication")

p_safety_allindications <- plot_grid(
    plot_safety_dots(toplot_safety_data_byindication %>% filter(indication_class == "All indications"))
    + ggtitle("All indications")
        + scale_x_discrete(labels = function(x) {
            stringr::str_wrap(x, width = 12)
        })
        + theme(plot.margin = margin(
            t = 5.5, b = 5.5, l = 5.5, r = -4, unit = "pt"
        )),
    plot_safety_numbers(toplot_safety_data_byindication %>% filter(indication_class == "All indications")) +
        theme(plot.margin = margin(
            t = 5.5, b = 5.5, l = 0, r = -4, unit = "pt"
        )),
    rel_widths = c(1, 0.9),
    align = "h",
    axis = "t",
    nrow = 1,
    ncol = 2
)
p_safety_onco <- plot_grid(
    plot_safety_dots(toplot_safety_data_byindication %>% filter(indication_class == "Oncology")) +
        ggtitle("Oncolology") +
        theme(
            plot.margin = margin(
                t = 5.5, b = 5.5, l = 0, r = -4, unit = "pt"
            ),
            axis.ticks = element_blank(),
            axis.title.y = element_blank(),
            axis.text.y = element_blank(),
            panel.spacing = unit(-0.1, "lines"),
            strip.background = element_blank(),
            axis.line = element_blank(),
            strip.text.y = element_blank()
        ),
    plot_safety_numbers(toplot_safety_data_byindication %>% filter(indication_class == "Oncology")) +
        theme(plot.margin = margin(
            t = 5.5, b = 5.5, l = 0, r = -4, unit = "pt"
        )),
    rel_widths = c(1, 1.12),
    align = "h",
    axis = "t",
    nrow = 1,
    ncol = 2
)
p_safety_nononco <- plot_grid(
    plot_safety_dots(toplot_safety_data_byindication %>% filter(indication_class == "Non-oncology")) +
        ggtitle("Non-oncolology") +
        theme(
            plot.margin = margin(
                t = 5.5, b = 5.5, l = 0, r = -4, unit = "pt"
            ),
            axis.ticks = element_blank(),
            axis.title.y = element_blank(),
            axis.text.y = element_blank(),
            panel.spacing = unit(-0.1, "lines"),
            strip.background = element_blank(),
            axis.line = element_blank(),
            strip.text.y = element_blank()
        ),
    plot_safety_numbers(toplot_safety_data_byindication %>% filter(indication_class == "Non-oncology")) +
        theme(plot.margin = margin(
            t = 5.5, b = 5.5, l = 0, r = 5.5, unit = "pt"
        )),
    rel_widths = c(1, 1.12),
    align = "h",
    axis = "t",
    nrow = 1,
    ncol = 2
)

p_safety_byindication <- plot_grid(
    p_safety_allindications + ggtitle("test"),
    p_safety_onco,
    p_safety_nononco,
    rel_widths = c(1, 0.83, 0.83),
    nrow = 1
)

p_safety_byindication
```